Enabling the Use of GPU Virtualization in Cloud Environments

### 2015-10-08 paper.tex
estructurado el state of art
Motivation improved

\section{\uppercase{State of Art}}
\label{sec:state}
Our solutions is based on GPU virtualization and sharing resources in order
to reach a fair balance between supply and demand.
Several attempts of achieving what we are pursing have been
made, but none of them are so ambitious as our project. On
the one hand the work done in~\cite{younge2013enabling} allows the VM managed
by the hypervisor Xen to access the GPUs in the physical
node, with the implied limitations of restricting the nodes not
to use more GPUs than the hosted in the machine and not
to share idle GPUs to other machines.
The solution presented by the project gVirtus~\cite{giunta2010gpgpu} virtualizes GPUs and make
them accessible for any VM in the cluster. However, this kind
of virtualization strongly depends on the hypervisor used, so
does its performance. Another similar solution is presented in~\cite{diab2013dynamic} by the name of gCloud. 
This one continues without being integrated in a Cloud Computing Manager, though its main
drawback is that the code of the applications must be modified
in order to be run in their virtual-GPU environment. A runtime component to provide abstraction
and sharing of GPUs is presented in~\cite{becchi2012virtual}, which allows scheduling policies 
to isolate and share GPUs in a cluster for a set of applications. Our proposal uses a different 
approach employing virtual machines and a cloud computing environment to allow to share GPUs in a flexible way.
On the other hand, the work exposed in~\cite{jungpgpu} is more mature, however,
it is only focussed on compute intensive HPC applications. 
The main idea of our proposal goes further and apart from
bringing solutions for all kind of HPC applications, it is aimed
to boost the cluster flexibility in the use of GPUs.

@@

Cloud computing, storage and other facilities are becoming 
more popular during the last years thanks to the easy set up 
process and the low required IT invest. The features provided 
by the cloud vendors are useful for companies and individuals 
which have access to the last technology without spending a 
huge amount of money. Moreover, the systems offered by the cloud 
companies matches with the ones required thanks to its flexibility.

The power of current cloud vendors is based on the use 
of Virtual Machines (VM) which  
emulates a physical node with its own 
CPUs, storage, network interconnection, OS and GPU. 
However, to provide a VM with GPUs means
that the guest host must be equipped with at least one GPU, and
once it has been assigned to an instance, no other VM can use
it. Therefore, the number of VM that are allowed to use a GPU is 
limited by the number of the installed devices. This lack of 
flexibility is also noticed in the cluster configurations, where 
depending on the users’ necessities, 
two approaches to introduce GPGPU computation can be determined:
the conservative, to have a lot of hosts with GPUs to be
ensured that you are likely to satisfy all the requests; or a
more daring approach, where you will count with some GPU-
enabled nodes hoping not to arrive a large burst of requests
so the cluster would run out of GPU resources.

The use of a virtualization technology such as {rCUDA} adds 
the desired flexibility because this software detaches the GPUs 
from the nodes where they are installed allowing all nodes in a cluster
not only use them as if they were local but also to share them 
achieving a higher utilization rate and reducing the idle status time.

Thanks to the shareability introduced by {rCUDA}, all VM in the cloud 
environment should be able to use and share all the GPUs installed in 
the physical nodes removing the current limitation. But, in order to achieve this goal, 
the cloud facilities need to be provided with the {rCUDA} software and 
in the case of OpenStack, a module for managing virtualized GPUs 
 has to be implemented.

****** FALTA UNIR AMBAS PARTES DEL TRABAJO ******

>>>>

\section{\uppercase{State of Art}}
\label{sec:state}
Our solutions is based on GPU virtualization and sharing resources in order
to reach a fair balance between supply and demand. 
Several attempts of achieving what we are pursing have been
made, but none of them are so ambitious as our project.

On
the one hand the work done in~\cite{younge2013enabling} allows the VM managed
by the hypervisor Xen to access the GPUs in the physical
node, with the implied limitations of restricting the nodes not
to use more GPUs than the hosted in the machine and not
to share idle GPUs to other machines. The solution presented 
by the project gVirtus~\cite{giunta2010gpgpu} virtualizes GPUs and make
them accessible for any VM in the cluster. However, this kind
of virtualization strongly depends on the hypervisor used, so
does its performance. Another similar solution is presented in~\cite{diab2013dynamic} by the name of gCloud. 
This one continues without being integrated in a Cloud Computing Manager, though its main
drawback is that the code of the applications must be modified
in order to be run in their virtual-GPU environment. A runtime component to provide abstraction
and sharing of GPUs is presented in~\cite{becchi2012virtual}, which allows scheduling policies 
to isolate and share GPUs in a cluster for a set of applications. Our proposal uses a different 
approach employing virtual machines and a cloud computing environment to allow to share GPUs in a flexible way.
On the other hand, the work exposed in~\cite{jungpgpu} is more mature, however,
it is only focussed on compute intensive HPC applications.

The main idea of our proposal goes further and apart from
bringing solutions for all kind of HPC applications, it is aimed
to boost the cluster flexibility in the use of GPUs.

@@

Cloud computing, storage and other facilities are becoming 
more popular during the last years thanks to the easy set up 
process and the low required IT invest. The features provided 
by the cloud services are useful for companies and individuals 
which have access to the last technology without spending a 
huge amount of money. Moreover, the systems offered by the cloud 
companies matches with the ones required thanks to its flexibility.

Two different kinds of cloud approaches are available:
on the one hand, public facilities such as Amazon Web Services or 
Microsoft Azure allows users to chose a by default system configuration 
or to compose its own system features. On the other hand, institutions/companies 
are able to deploy their own cloud infrastructure by using software products 
like OpenStack.

In both cases, the power of current cloud facilities is based on the use 
of Virtual Machines (VM) which  
emulates a physical node with its own 
CPUs, storage, network interconnection, OS and GPU. 
However, to provide a VM with GPUs means
that the guest host must be equipped with at least one GPU, and
once it has been assigned to an instance, no other VM can use
it. Therefore, the number of VM that are allowed to use GPUs is 
limited by the number of the installed devices. This lack of 
flexibility is also noticed in the cluster configurations, where 
depending on the users’ necessities, 
two approaches to introduce GPGPU computation can be determined:
the conservative, to have a lot of hosts with GPUs to be
ensured that you are likely to satisfy all the requests; or a
more daring approach, where you will count with some GPU-
enabled nodes hoping not to arrive a large burst of requests
so the cluster would run out of GPU resources.

The use of a virtualization technology such as {rCUDA} adds 
the desired flexibility because this software detaches the GPUs 
from the nodes where they are installed allowing all nodes in a cluster
not only use them as if they were local but also to share them 
achieving a higher utilization rate and reducing the idle status time.
Thanks to the shareability introduced by {rCUDA}, all VM in the cloud 
environment should be able to use and share all the GPUs installed in 
the physical nodes removing the current limitations.

The main goals of this work are to study current cloud solutions from 
the point of view of HPC field and GPU usage, and to analyze and improve 
them by adding flexibility with GPU virtualization.

Our work have been focused on Amazon Web Service as example of public 
solution and on OpenStack software acting as a private cloud environment. 
In order to achieve it, these cloud facilities need to be provided with the {rCUDA} software and 
in the case of OpenStack, a module for managing virtualized GPUs 
 requires to be implemented.
